---
title: "Embryonic area in AT timed matings"
author: "Andrew Garcia"
date: "2024-04-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
embryos <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRardz8DQxyjal-BQCrVOqgcEo2Rug6MRojf1h-yJxMPz6NCxl1SqHNmSZx44qS1AYn3_LwYYCtAvyj/pub?gid=260011419&single=true&output=csv")

head(embryos)
embryos$Naa10.genotype <- as.factor(embryos$Naa10.genotype)
embryos$Litter.Theiler.stage <-as.factor(embryos$Litter.Theiler.stage)
structure(embryos$area)
```


```{r}
install.packages('tidyverse', repos = "http://cran.us.r-project.org")
```
```{r}
install.packages('ggplot2',repos = "http://cran.us.r-project.org")
```



```{r}
install.packages('Rmisc',repos = "http://cran.us.r-project.org")
```


```{r}
library(Rmisc)
area.summ <- summarySE(data = embryos, "area..um.",
          groupvars = c("Naa10.genotype","Litter.Theiler.stage"))

area.summ
```

```{r}
install.packages('car', repos = "http://cran.us.r-project.org")
library(car)
dev.lm <- lm(area..um. ~ Litter.Theiler.stage+Naa10.genotype,embryos)
summary(dev.lm)
```

```{r}
library(car)
Anova(dev.lm, type = "III")
```

```{r}
install.packages('multcomp', repos = "http://cran.us.r-project.org")
```


```{r}
library(multcomp)

dev.tukey <- glht(dev.lm,linfct = mcp(Naa10.genotype = "Tukey"))
summary(dev.tukey)

```

```{r}
library(ggplot2)
emb.dev <- ggplot(embryos, aes(x= Litter.Theiler.stage,y=area..um.,shape = Naa10.genotype))+
  geom_point(aes(shape = Naa10.genotype))+
     geom_boxplot(outlier.shape = NA,
               alpha = 0.1)+
    stat_summary(fun = mean, color = "black", position = position_dodge(0.75),width = 0.2,
             geom = "crossbar",
             show.legend = FALSE)+
   geom_errorbar(data = area.summ, aes(ymin=area..um.-sd,
                     ymax=area..um.+sd),
                 width=.2,
                 position=position_dodge(0.75))+
  xlab("Theiler stage of litter")+
  ylab("area (um^2)")+
  ggtitle("Embryo area during development")

emb.dev
  
```

Next (2024-04-23 11:58 AM):
1) Finish measuring remaining embryos

2) May need to omit females because of insufficient numbers (delayed development more sporadic in females?)

3) Can do two-sample t-tests if I plot Theiler stages individually or ANOVA if I want to have all Theiler stages on plot

Updates (2024-04-29 11:55 AM)
1) Find matching number of controls for TS15 Naa10 KO/Y embryos probably from another litter
2) ID 1-2 TS16 Naa10 KO/Y embryos if needed.
3) May need another column variable (e.g., delayed or normal developing)
4) May bucket TS12-TS13. Or pair TS14-15, TS16-17, TS17-18

Only one TS16 litter, so may need to bucket with TS17 and thus bucket as described
2024-04-29 item 4.

Naa10 KO/Y embryos may indeed be delayed developing, but also intrinsically smaller.
