---
title: "AT embryo delayed development measurements"
output:
  html_document:
    df_print: paged
Author: Andrew Garcia
---

# Embryonic perimeter in males only

Imported only the male embryo measurements because the female sample size
is much smaller. Delay observed in females, but occurs more sporadically.

```{r}
embryos <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRardz8DQxyjal-BQCrVOqgcEo2Rug6MRojf1h-yJxMPz6NCxl1SqHNmSZx44qS1AYn3_LwYYCtAvyj/pub?gid=260011419&single=true&output=csv")

head(embryos)
embryos$Naa10.genotype <- as.factor(embryos$Naa10.genotype)
embryos$Litter.Theiler.stage <-as.factor(embryos$Litter.Theiler.stage)
structure(embryos$area)
```


```{r}
install.packages("tidyverse",repos='http://cran.us.r-project.org')
install.packages("ggplot2",repos='http://cran.us.r-project.org')
```


```{r}
install.packages("Rmisc",repos='http://cran.us.r-project.org')
```

## Perimeter analysis
```{r}
library(Rmisc)
peri.summ <- summarySE(data = embryos, "perimeter..mm.",
          groupvars = c("Naa10.genotype","Litter.Theiler.stage"))

peri.summ
```

```{r}
install.packages("car",repos='http://cran.us.r-project.org')
library(car)
peri.lm <- lm(perimeter..mm. ~ Litter.Theiler.stage+Naa10.genotype,embryos)
summary(peri.lm)
```

```{r}
library(car)
Anova(peri.lm, type = "III")
```

```{r}
plot(peri.lm)
```

```{r}
install.packages("multcomp",repos='http://cran.us.r-project.org')
```


```{r}
library(multcomp)

dev.tukey <- glht(peri.lm,linfct = mcp(Naa10.genotype = "Tukey"))
summary(dev.tukey)

ts.tukey <- glht(peri.lm,linfct = mcp(Litter.Theiler.stage = "Tukey"))
summary(ts.tukey)
```

```{r}
library(ggplot2)
peri.plot <- ggplot(embryos, aes(x= Litter.Theiler.stage,y=perimeter..mm.,shape = Naa10.genotype))+
  geom_point(aes(shape = Naa10.genotype))+
     geom_boxplot(outlier.shape = NA,
               alpha = 0.1)+
    stat_summary(fun = mean, color = "black", position = position_dodge(0.75),width = 0.2,
             geom = "crossbar",
             show.legend = FALSE)+
   geom_errorbar(data = peri.summ, aes(ymin=perimeter..mm.-sd,
                     ymax=perimeter..mm.+sd),
                 width=.2,
                 position=position_dodge(0.75))+
  xlab("Theiler stage of litter")+
  ylab("perimeter (mm)")+
  ggtitle("Embryo perimeter during development")+
  theme_bw(base_size = 12, base_family = "Helvetica")

peri.plot
  
```


## Area analysis

```{r}
area.peri.plot <- ggplot(embryos, aes(x= perimeter..mm.,
                    y=area..mm.2.,
                    shape = Naa10.genotype))+
  geom_point(aes(shape = Naa10.genotype))

area.peri.plot
```


```{r}
library(Rmisc)
area.summ <- summarySE(data = embryos, "area..mm.2.",
          groupvars = c("Naa10.genotype","Litter.Theiler.stage"))

area.summ
```

```{r}
library(car)
area.lm <- lm(area..mm.2. ~ Litter.Theiler.stage+Naa10.genotype,embryos)
summary(area.lm)
```


```{r}
plot(area.lm)
```

```{r}
library(car)
Anova(area.lm, type = "III")
```



```{r}
library(multcomp)

dev.tukey <- glht(area.lm,linfct = mcp(Naa10.genotype = "Tukey"))
summary(dev.tukey)

```

```{r}
library(ggplot2)
area.plot <- ggplot(embryos, aes(x= Litter.Theiler.stage,y=area..mm.2.,shape = Naa10.genotype))+
  geom_point(aes(shape = Naa10.genotype))+
     geom_boxplot(outlier.shape = NA,
               alpha = 0.1)+
    stat_summary(fun = mean, color = "black", position = position_dodge(0.75),width = 0.2,
             geom = "crossbar",
             show.legend = FALSE)+
   geom_errorbar(data = area.summ, aes(ymin=area..mm.2.-sd,
                     ymax=area..mm.2.+sd),
                 width=.2,
                 position=position_dodge(0.75))+
  xlab("Theiler stage of litter")+
  ylab("area (mm^2)")+
  ggtitle("embryo area during development
          (n=32)")+
  theme_bw(base_size = 12, base_family = "Helvetica")

area.plot
```



## alternative model?

```{r}
library(car)
alt.lm <- lm(area..mm.2. ~ perimeter..mm.*Litter.Theiler.stage*Naa10.genotype,embryos)
summary(area.lm)
```


```{r}
library(ggplot2)
litter.plot <- ggplot(embryos, aes(x= Litter,y=perimeter..mm.,shape = Naa10.genotype))+
  geom_point(aes(shape = Naa10.genotype))+
     geom_boxplot(outlier.shape = NA,
               alpha = 0.1)+
  xlab("Theiler stage of litter")+
  ylab("perimeter (mm)")+
  ggtitle("Embryo perimeter during development
          (n=32)")+
  theme_bw(base_size = 12, base_family = "Helvetica")

litter.plot
```


Next (2024-04-23 11:58 AM):
1) Finish measuring remaining embryos

2) May need to omit females because of insufficient numbers (delayed development more sporadic in females?)

3) Can do two-sample t-tests if I plot Theiler stages individually or ANOVA if I want to have all Theiler stages on plot

Updates (2024-04-29 11:55 AM)
1) Find matching number of controls for TS15 Naa10 KO/Y embryos probably from another litter
2) ID 1-2 TS16 Naa10 KO/Y embryos if needed.
3) May need another column variable (e.g., delayed or normal developing)
4) May bucket TS12-TS13. Or pair TS14-15, TS16-17, TS17-18

Only one TS16 litter, so may need to bucket with TS17 and thus bucket as described
2024-04-29 item 4.

Naa10 KO/Y embryos may indeed be delayed developing, but also intrinsically smaller.
